\clearpage
\section{Conclusion}
\label{sec:conclusion}

This thesis set out to address a fundamental tension in confined-space drone operations: when collisions with the environment are inevitable, how should the platform be designed to survive them, and can the onboard sensor signatures of those collisions be exploited for spatial awareness? The work followed two parallel threads --- mechanical deflection and inertial inference --- and the results from each thread reinforce a common theme: the rotating cage mechanism fundamentally redistributes collision energy, with measurable and operationally meaningful consequences for both survivability and sensor interpretation.

\subsection{Summary of Contributions}

\textbf{Platform Design.} A custom 4-inch quadcopter was designed, built, and flight-validated. The platform's dimensions were governed by a three-way trade-off: a lower bound set by the payload requirements of the Raspberry~Pi~5 companion computer and PETG protective cage, and an upper bound set by indoor safety limits on propeller size and kinetic energy. A 6S power system paired with 2004-stator motors provides a thrust-to-weight ratio of approximately 3.8:1, placing hover at 45--50\% throttle and reserving substantial headroom for post-collision recovery. The autonomous flight stack integrates PX4 firmware, ROS~2 middleware, and an OptiTrack MoCap ground-truth system through the EKF2 estimator, with the \texttt{flight\_director.py} mission executor driving a reproducible collision-loop state machine. The PETG spherical cage (35.8\,cm diameter) was tested in two configurations --- rigidly mounted (Fixed Cage) and bearing-decoupled (Rotating Cage) --- isolating the bearing mechanism as the single experimental independent variable.

\textbf{Mechanical Performance.} Across 127 verified structural impacts (67 Fixed, 60 Rotating), the rotating cage reduced maximum deceleration by 59\% relative to the fixed configuration ($-1.64 \pm 0.64$\,m/s$^2$ vs.\ $-1.03 \pm 0.50$\,m/s$^2$, $p < 0.001$, $d = 1.05$). IMU vibration analysis revealed qualitatively distinct post-impact behaviors: the fixed cage sustained high-frequency pitch-and-roll oscillations consistent with stick-slip contact dynamics, while the rotating cage's bearing decoupling allowed the internal IMU to return to a nominal noise floor substantially faster. The aggregate IMU envelope (mean~$\pm\sigma$ across all 127 flights) confirmed a narrower acceleration-deviation spread for the rotating configuration.

The deflection benefit carries a modest spatial cost. The rotating cage exhibited a slightly wider maximum trajectory deviation ($131.0 \pm 32.1$\,mm vs.\ $118.4 \pm 30.4$\,mm, $p = 0.022$, $d = 0.41$), consistent with the bearing mechanism transferring angular momentum at the moment of contact. However, the total recovery area --- the Spatial Integral of Absolute Error --- remained statistically indistinguishable between configurations ($p = 0.998$), indicating that the flight controller's PID correction capacity absorbs this offset without expanding the overall spatial footprint of the recovery maneuver.

\textbf{Impact-Angle Inference.} A Random Forest regressor trained on 14 IMU-derived features recovered impact angle with $R^2 \approx 0.72$ and mean absolute error of $4.5^\circ$--$4.9^\circ$ for both cage configurations. This level of accuracy supports rudimentary tactile awareness --- the drone can distinguish a shallow glancing impact ($\sim$30$^\circ$) from a deeper collision ($\sim$60$^\circ$) using only the vibration signature recorded by its onboard IMU.

Three subsidiary findings carry practical significance. First, the rotating cage's angle signature is invisible to any linear model (Huber baseline $R^2 = 0.004$) yet fully recovered by the non-linear Random Forest ($R^2 = 0.719$), demonstrating that the bearing mechanism disperses angle information across a distributed multivariate fingerprint rather than destroying it. Second, cross-condition transfer fails decisively in both directions ($R^2 < 0.26$), meaning a model trained on one cage type cannot be deployed on the other without a mechanical-normalisation transfer function. Third, persistent learning-curve gaps of approximately 0.10 $R^2$ between training and validation performance indicate that the current dataset of 127 flights is insufficient to fully saturate the 14-dimensional feature space, and that additional data collection or stronger regularisation would close this gap.

\subsection{Limitations}

Several constraints bound the generalizability of these findings. The experiments were conducted at two approach angles ($45^\circ$ and $75^\circ$) against a single cylindrical obstacle with a cardboard surface. Performance at acute angles, against flat walls, corners, or compliant surfaces, and on materials with different friction coefficients may diverge from the patterns observed here. The collision target was static; real-world confined-space navigation may involve moving or yielding obstacles. All trajectory analysis relied on post-processed MoCap ground truth at 120\,Hz; a fully onboard system operating without external tracking would need to substitute this with visual-inertial odometry or pure inertial propagation, each of which introduces its own drift and latency characteristics.

The Random Forest model captures non-linear structure in the IMU-to-angle mapping but does not model temporal dynamics explicitly. A recurrent or attention-based architecture that ingests the full collision time series rather than aggregated window statistics could potentially close the residual non-linearity visible in the residual plots.

\subsection{Future Work}

Three directions emerge naturally from the findings.

\textbf{Expanding the collision envelope.} The experimental matrix should be extended to additional obstacle geometries (flat walls, convex corners, pipes of varying diameter), surface materials (metal, rubber, glass), and approach angles spanning the full $0^\circ$--$90^\circ$ range. Multi-impact sequences --- where the drone ricochets between surfaces before recovering --- would test whether the vibration-decoupling benefit of the rotating cage compounds across successive contacts.

\textbf{Closing the sim-to-real and lab-to-field gap.} The MoCap ground-truth dependency should be progressively relaxed. A natural intermediate step is to replace the external tracking system with onboard visual-inertial odometry (VIO) and compare the resulting velocity estimates against the EKF2 states used in this work. Beyond the laboratory, testing in a realistic industrial mock-up (e.g., a stainless steel tank interior with pipe networks) would expose the platform to the multi-path radio reflections, ferromagnetic compass interference, and feature-poor visual surfaces that characterize genuine confined-space environments.

\textbf{Toward tactile SLAM.} The demonstration that IMU signatures alone recover impact angle with $\sim$5$^\circ$ accuracy opens the door to a broader research program: \emph{tactile simultaneous localization and mapping}. If a drone can estimate not only \emph{that} it struck an obstacle but \emph{at what angle} and with \emph{what surface normal}, successive impacts against the same surface could be triangulated to build a coarse geometric map of the environment purely from contact signatures. Combined with the mechanical resilience of the rotating cage, this would realize the paradigm shift articulated in the introduction: from a drone that is blind but survivable to one that is blind but tactile --- navigating by touch in spaces where no other sensor modality is viable.
