\section{Related work}
\subsection{Collision-Resilient Drone Designs for Confined spaces}
\textit{Here mention the congifly \cite{azambujaWhenBeingSoft2022}, spinfly\cite{lercheSpinFlyUAVSystem} as something I mainly continued and learned from , Flyability\cite{UltimateGuideDrone} cluttered environmnet inspiration to how to build the cage \cite{mahmudDevelopmentCollisionResilient2021} talk about gomboc and what not
FOR NEXT TIME -> start easy on the first 2 things, perhaps talk about th edevelopment of the drone first in the Methodology section}

This paper is a spiritual continuation of Spinfly \citep{lercheSpinFlyUAVSystem} by A. Lerche. 
Lerche had a a similiar problem statement and he tried to develop a drone for autonomous inspection, in particular odometry. From his work and personal conversations with him a number of lessons to avoid in this work has been identified. Lerche used iNAV flight control framework as backbone of his drone development. While iNAV allows for autonomous flight it si not particularly well suited for fast development. Lerche had spent a considerable time and resources on hardware connections, particularly soldering of Flight Controller and other devices together and various troubleshooting of hardware and software in iNAV. 
Constraint of navigating cluttered spaces safely brought me to work on Robust Aerial Swarms by  \cite{mulgaonkarRobustAerialRobot2018} where instead of active obstacles detection and avoidance they instead crashed drones and have them recover from the crash, albeit with drone swarms each weighing 25grams. This takes a radically different approach to solving 'how to get from A to B' as   Simultaneous localization and visualization (SLAM) is at a forefront of robotics and in many ways its holygrail, Mulgaonkar has instead chosen to map environment by clashing with it, and instead vicariously map it by localization of the drones themselves. And instead of relaying on detecting obstacle first and never clashing with it, have resilient enough drone which can instead recover from the clash and continue its mission. As SLAM is being developed it is not a ready technology for relatively small drones. Even best automakers in the world have still not brought this technology to navigate reliably any environment at the moment with state of the art LIDARs and other expensive and somewhat heavy equipment. While unmapped and changing environment of potential drone use in cluttered environment can be many fold more difficult to navigate and thus require quite sizable, heavy and expensive equipment if the drone was to go down the SLAM road for autonomy. 
While small drones in the aerial swarms could navigate their environment with clashing, using a a simple cage, the drone proposed in this work should be able of partially autonomous flying when being denied signal and even have ability to inspect its environment with various sensors it might use for it. This warrants some functional payload overhead which due to lift physics can be overcome only so much by more powerful motors, and at certain point the drone must grow in size. For this project a 4inch propeller drone design was chosen based on criteria mentioned later. At this size the penalty of hitting something increases substantially \textit{DATA???? citetions? soemthing to proove this claim???} as the ratio of inertia and relative structural strenght of the drone and its propellers decreases, and on the other hand with the increased size of the drone everything in its environment is relatively smaller. Meaning that while Mulgaonkar could rely on simple 2 tube cage spanning their drones to avoid most items in their environment, the protection of a 4inch drone (4inch propellers -> at minimum  20cm in X and Y dimmensions) must be such that this drone is protected from corners of items such as shelves or continuous pipping and that such objects do not directly interfere with the drone, and with its propellers in particular. Here a research \citealp[Development of Collision Resilient Drone for Flying in Cluttered Environment]{mahmudDevelopmentCollisionResilient2021} helps to tackle this issue. He proposes either a gimbal like sphere which freely spins upon contact with its environment, similiar as a commercial product by Flaybility \cite{UltimateGuideDrone} \textit{Sourcing missing, fucked up ZOTERO} and a Gombok \cite{Gomboc2026} \textit{eve fucking wikipedia reference will do here}. Gombok shape, or in this case a turtle-like shape has an interesting property of being potentially stable only in one orientation, when a turtle is turned upside down it can just by its own movement come back to its feet thanks to its center of gravity being suddenly high up in such case, making it more unstable. Mahmout has laveraged these properties in their design as well and concluded his paper with it being the superior cage design of the two. A specific use case given by supervisor of this paper comes to mind: inspection of stainless steel tanks. Today this is executed by a human who needs to descent into the tank through a menhole on the top. Hypothetically a turle shaped drone cage, which would after inspecting the tank ascend to fly out through the menhole, could more easily get through the narrow opening even if it would not navigate perfectly out of the hole, as if only the center of the drone would be within the manhole limits the drone could somewhat stably ascend without it being rocked too hard, as it would leverage its shape and its low center of gravity. Due to this hypothesis, the resulsts Mahmut had with their cage design, relative simplicity opposed to gimbal sphere and even possibility of using it similarly as Lerche did for odometry, the spinning turtle shape was eventually chosen for the drone design. 

Apart from physical survivability of larger drones in cluttered environments a topic of software control for autonomous flight had to be addressed. While this is a frontier technological development topic, 2 autonomous flight platforms standout with their rate of adoption and technological maturity, ArduPilot and PX4. Both are widely adopted and have a lively base of developers who add to their codebases. Apart from more free licensing of PX4 (which even allows for commercial extension based on the PX4 for proprietory purposes, while Ardupilot requires all development to be added to their codebase CITATION MISSING AGAIN) PX4 differs from Ardupilot having more modularized architecture and having more well developed hardware standards. This gave PX4 the edge in this work as it allowed for more straightforward attitude to the development and using standardized parts presumably with no or little troubleshooting.  